---
title: "MVA-Exploratory Data Analysis (EDA): Visualization"
output: html_document
date: "2024-02-16"
---

#Questions 

1) What are the common loan amounts that the borrower's request?

2) Highest frequency of Loan grade?

3) Variation of loan amounts requested by borrowers across different loan grades?
4) How does the observed variation in loan amounts requested by borrowers across different loan grades contribute to the creditworthiness?
5) Do higher loan amounts correspond to higher or lower loan percent incomes?
6) Common loan amount and loan percent income combinations within the dataset.

```{r}
library(readr)
dataset<- read_csv("C:/Users/rusha/Downloads/Multivariate/credit_risk_dataset.csv/credit_risk_dataset.csv")

```
```{r}
str(dataset)
```

```{r}
boxplot(dataset$person_income)
boxplot(dataset[,4:8])
boxplot(dataset$loan_percent_income)
```

#Insights:

Boxplot of person income and loan percent income shows an outlier.
The median for loan amount is close to 24000, with bottom edge of 10000 and top edge of 25000.
The median for loan percent income is close to 0.25, with bottom edge of 0.1 and top edge of 0.35.

```{r}
stars(dataset,labels = dataset$loan_status)
```

#Correlation matrix

```{r}
library(ggcorrplot)
correlation_matrix<- cor(dataset[,0:8])
ggcorrplot(correlation_matrix,type = "lower", lab = TRUE)

```

#Insights:

By looking at the correlation matrix we can understand the magnitude and direction of relationship between two variables.Closer the value to 1 the stronger the relationship. Negative value indicates that the direction is opposite. 

1) Here, value -0.65 for person_income vs loan_percent_income indicates a moderate-strong relationship and an inverse relationship.One variable increases, the other variable decreases, and vice versa.

2) Value 0.71 for loan_status vs loan_percent_income indicates a strong relationship and a positive correlation.
3) Value 0.38 for loan_amnt vs loan_percent_income indicates a moderate-strong relationship and a positive correlation. 
4) Value 0.25 for loan_amnt vs person_income indicates a weak relationship and a positive correlation. 


```{r}
# MAX AND MIN AGE
max_age <- max(dataset$person_age, na.rm = TRUE)
min_age <- min(dataset$person_age, na.rm = TRUE)
cat("Maximum Age:", max_age, "\n")
cat("Minimum Age:", min_age, "\n")

# People with an age between x and y
age_group <- function(arr, dataset) {
  for (i in seq_along(arr[-1])) {
    age_min <- arr[i] + 1
    age_max <- arr[i + 1]
    num_people <- sum(dataset$person_age >= age_min & dataset$person_age < age_max, na.rm = TRUE)
    cat("Age between", age_min, "and", age_max, ": Number of people", num_people, "\n")
  }
}

age_group(c(0, 18, 26, 36, 46, 56, 66), dataset)


```


# Univariate visualizations
1) Density plot of Loan Amount

```{r}
library(scales)
density_loan_amount <- ggplot(dataset, aes(x = loan_amnt)) +
  geom_density(fill = "grey") + labs(title = "Density Plot of Loan Amount") +
  scale_y_continuous(labels = comma)+
  xlab("Loan Amount") + ylab("Density")+theme_classic()

print(density_loan_amount)
```

#Insights:
The plot is skewed to the right. The peak of the density plot is somewhere around 25000. 
The plot is continuous and widely spread. A wider spread suggests a greater range of loan amounts. 


2) Bar plot of Loan Grade

```{r}
bar_loan_grade <- ggplot(dataset, aes(x = loan_grade)) +
  geom_bar(fill = "lightgreen") +
  labs(title = "Bar Plot of Loan Grade") +
  xlab("Loan Grade") +
  ylab("Frequency")+theme_bw()

print(bar_loan_grade)
```

#Insights:

Looking at the bar plot, we can understand Grade B is the highest in frequency while grade G is the least. 
Further, this means that the number of borrowers who are relatively low-risk, but not as creditworthy as Grade A are the most in count. 

The borrower's creditworthiness is moderate, that is grade C is more than the borrowers, grade A, who has the highest creditworthiness.


#Bivariate Visuals

1)Scatter Plot- Matrix

```{r}
library(GGally)
ggscatmat(dataset, columns=3:6, color="loan_status")
```




2) Boxplot of Loan grade vs Loan amount

```{r}
boxplot_loan_grade_amount <- ggplot(dataset, aes(x = loan_grade, y = loan_amnt)) +
  geom_boxplot(fill = "lightgreen") + 
  labs(title = "Box Plot of Loan Amount by Loan Grade") +
  xlab("Loan Grade") +
  ylab("Loan Amount") + theme_bw()

print(boxplot_loan_grade_amount)
```

#Insights:
The plot shows, that the borrowers with loan grade D has most number of outliers when compared with the loan amount requested by them. This might also define why the borrower's is considered to have higher risk of creditworthiness. 



3) Scatter plot of Loan amount vs loan percent income

```{r}
scatter_plot <- ggplot(dataset, aes(x = loan_amnt, y = loan_percent_income )) +
  geom_point() +  # Add points
  labs(x = "Loan Amount", y = "Loan Percent Income", title = "Scatter Plot of Loan Amount vs Loan Percent Income")
plot(scatter_plot)
```

Scatter Plot of Loan amount vs loan percent income with loan grades colored 

```{r}
ggplot(dataset, aes(x = loan_amnt, y = loan_percent_income, color = loan_grade)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Scatter plot of Loan Amount vs. Loan Percent Income by Loan Grade",
       x = "Loan Amount", y = "Loan Percent Income", color = "Loan Grade") +
  theme_minimal()
```

#Insights:

These is no linear relationship between loan amount and loan percent income.

Points are more densely populated between loan amount 0-10000 and 22000-30000.

Having a high loan percent income does not always corresponds to having a high loan amount. Borrowers who have requested Loan amounts of 5000-7000 also corresponds to high percentage of borrowers income. The same for borrowers requesting loan about of 25000 or greater. 



```{r}
library(lattice)
xyplot(person_income ~ person_age | loan_grade, data = dataset)
```


