---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

1) Model development
```{r}
library(MASS)
library(ggplot2)
library(memisc)
library(ROCR)
library(dplyr)
library(klaR)

data <- read.csv("C:/Users/rusha/Downloads/Multivariate/credit_risk_dataset.csv/credit_risk_dataset.csv")
str(data)
data$person_income <- as.numeric(data$person_income)
data$loan_amnt <- as.numeric(data$loan_amnt)


# Define the breaks for the income ranges
breaks <- c(0, 25000, 50000, 80000, 120000, 160000, Inf)

# Define labels for the income ranges
labels <- c("0-25000", "25001-50000", "50001-80000", "80001-120000", "120001-160000", "160001+")

# Discretize the income values into bins
data$person_income <- cut(data$person_income, breaks = breaks, labels = labels)

# Check the result
head(data$person_income)

str(data)
```

```{r}
#wdbc.data <- as.matrix(wdbc[,c(3:32)])
#row.names(wdbc.data) <- wdbc$id
#wdbc_raw <- cbind(wdbc.data, as.numeric(as.factor(wdbc$diagnosis))-1)
#smp_size_raw <- floor(0.75 * nrow(wdbc_raw))
#train_ind_raw <- sample(nrow(wdbc_raw), size = smp_size_raw)
#train_raw.df <- as.data.frame(wdbc_raw[train_ind_raw, ])
#test_raw.df <- as.data.frame(wdbc_raw[-train_ind_raw, ])
#wdbc_raw.lda <- lda(formula = train_raw.df$diagnosis ~ ., data = train_raw.df)
#wdbc_raw.lda

# Combine loan data with loan status and income group
loan.data <- as.matrix(data[,c(6:7)])
loan.data <- cbind(loan.data, loan_status = data$loan_status)
row.names(loan.data) <- data$loan_status
loan_raw <- cbind(loan.data, as.numeric(as.factor(data$person_income))-1)
colnames(loan_raw)[3] <- "income_group"
# Remove rows with NA values
loan_raw <- loan_raw[complete.cases(loan_raw), ]
# Split data into train and test sets
samp_size_raw <- floor(0.75 * nrow(loan_raw))
train_loan_raw <- sample(nrow(loan_raw), size = samp_size_raw)
train_raw.df <- as.data.frame(loan_raw[train_loan_raw, ])
test_raw.df <- as.data.frame(loan_raw[-train_loan_raw, ])
# Fit LDA model
loan_raw.lda <- lda(formula = train_raw.df$income_group ~., data = train_raw.df)
loan_raw.lda
```

Insights
The prior probability of group 0 is approximately 0.43, and the prior probability of group 1 is approximately 0.57, indicating that the training dataset is slightly imbalanced, with more observations belonging to group 1.(i.e loan accepted)

loan_int_rate and loan_percent_income have positive coefficients, suggesting that higher values of these variables are associated with group 1.
'V4' or 'income_group' has a negative coefficient, indicating that higher values of V4 are associated with group 0.

```{r}
plot(loan_raw.lda, xlab = "Predicted Group")
```


```{r}
#wdbc_raw.lda.predict <- predict(wdbc_raw.lda, newdata = test_raw.df)
#wdbc_raw.lda.predict$class
#wdbc_raw.lda.predict$x

#loan_raw.lda.predict <- predict(loan_raw.lda, newdata = test_raw.df)
#loan_raw.lda.predict$class
#loan_raw.lda.predict$x

# Remove NA values from the data frame
clean_test_raw.df <- na.omit(test_raw.df)

# Perform prediction with LDA
loan_raw.lda.predict <- predict(loan_raw.lda, newdata = clean_test_raw.df)

# Access predicted class
loan_raw.lda.predict$class

# Access LD scores
loan_raw.lda.predict$x
```

```{r}
# Get the posteriors as a dataframe.
loan_raw.lda.predict.posteriors <- as.data.frame(loan_raw.lda.predict$posterior)

pred <- prediction(loan_raw.lda.predict.posteriors[,2], test_raw.df$income_group)
roc.perf = performance(pred, measure = "tpr", x.measure = "fpr")
auc.train <- performance(pred, measure = "auc")
auc.train <- auc.train@y.values
plot(roc.perf)
abline(a=0, b= 1)
text(x = .25, y = .65 ,paste("AUC = ", round(auc.train[[1]],3), sep = ""))

# The AUC value will be different everytime we compute the code as we are taking random sample for test and train model.
```

