---
title: "Mean and variance analysis"
output: html_document
date: "2024-02-23"
---

```{r}
library(readr)
raw_dataset<- read_csv("C:/Users/rusha/Downloads/Multivariate/credit_risk_dataset.csv/credit_risk_dataset.csv")

```
```{r}
str(raw_dataset)
```
Loading the dataset. 

```{r}
options(scipen = 999) # To suppress scientific notation 

#dataset <- raw_dataset[,3:8]
dataset_clean <- na.omit(raw_dataset)
indvar_dataset <- dataset_clean[,2:8]
```


Univariate analysis 
1) Finding means of each column
2) Covariance matrix for all columns in the data
3) Mahalanobis distances for each observation: To identify outlier or observations that are unusual relative to the overall distribution of the data. The distance is also used in clustering algorithms, like k-means clustering. 


```{r}
dataset_cm <- colMeans(indvar_dataset, na.rm = TRUE)
dataset_cm

dataset_S <- cov(indvar_dataset)
dataset_S

dataset_d <- apply(indvar_dataset, MARGIN = 1, function(indvar_dataset)t(indvar_dataset - dataset_cm) %*% solve(dataset_S) %*% (indvar_dataset - dataset_cm))
dataset_d
```


Insights: 
Correlation
Average Income: The average person's income is approximately $104,516.41.
Average Loan Amount: The average loan amount is approximately $19,726.53.
Loan-to-Income Ratio: On average, loans represent about 25.63% of a person's income.

Covariance between person_income and loan_amnt is indicates strong relationship between them. 

Distance 
Large distance = outlier / anomaly


Lets find out if the dataset follows a particular distribution/ is normally distributed?
--Applying qqnorm to three variables: person_income, loan_amnt, and loan_percent_income
```{r}
qqnorm(dataset_clean$person_income, main = "Person income")
qqline(dataset_clean$person_income)

qqnorm(dataset_clean$loan_amnt, main = "Loan amount")
qqline(dataset_clean$loan_amnt)

qqnorm(dataset_clean$loan_percent_income, main = "Loan to income ratio")
qqline(dataset_clean$loan_percent_income)
```

Insight: 
Loan amount is not normally distributed
Person income is following the reference line at some points, while misses at some. 
Loan to income ratio is normally distributed for most points at the centre, while misses some points at the start and end. 


Test to find equality of variances between 2 groups of loan status. 
```{r}
var.test(raw_dataset$loan_status==0, raw_dataset$loan_status==1)
```
Insight: 
p-value = 1, suggests that there is no significant difference in the variances between the two groups of loan status. 



Performing t-test for each independent variable in the dataset, by comparing its means between two groups of loan status. 

```{r}
with(data=dataset_clean,t.test(person_income[loan_status==0],person_income[loan_status==1],var.equal=TRUE))
```

```{r}
with(data=dataset_clean,t.test(person_emp_length[loan_status==0],person_emp_length[loan_status==1],var.equal=TRUE))
```

```{r}
with(data=dataset_clean,t.test(loan_amnt[loan_status==0],loan_amnt[loan_status==1],var.equal=TRUE))
```

```{r}
with(data=dataset_clean,t.test(loan_int_rate[loan_status==0],loan_int_rate[loan_status==1],var.equal=TRUE))
```

```{r}
with(data=dataset_clean,t.test(loan_percent_income[loan_status==0],loan_percent_income[loan_status==1],var.equal=TRUE))
```
```{r}
with(data=dataset_clean,t.test(cb_person_cred_hist_length[loan_status==0],cb_person_cred_hist_length[loan_status==1],var.equal=TRUE))
```


Insight: 
p-value < 0.00000000000000022 (Shows the strongest evidence against the null hypothesis)
p-value = 0.000009249 (Shows strong)
That is reject the null hypothesis.

p-value = 0.1021 (Fail to reject or accept the null hypothesis)




Multivariate analysis: 
Hotelling's T-squared test/ extension of the univariate t-test to multidimensional data.


```{r}
library(Hotelling)

 

t2test <- hotelling.test(person_income + person_emp_length + loan_amnt + loan_int_rate + loan_percent_income + cb_person_cred_hist_length ~ loan_status, data = dataset_clean )
cat("T2 statistic =", t2test$stat[[1]],"\n")
print(t2test)
```
Insight: 
T2 statistic of 645, measures how much the average values of multiple variables differ between the two groups in loan status. Larger T2 value indicates a greater difference between the group means in the multivariate space. In simple terms, stronger distinction between the groups based on the combination of variables being analyzed. 

p-value = 0 suggests strong evidence against the null hypothesis. 
Thus indicating significant differences between the group means in the multivariate space. That is there is something affecting and we cannot accept the null hypothesis.



Levene's test--to find if the variance of independent variables differ significantly between different groups of loan status. 
```{r}
datastand <- scale(dataset_clean[,3:8])
datastand
```

```{r}
attach(dataset_clean)
datanodef <- datastand[loan_status==0,]
datanodef
```


Vector containing the median values of each column in the subset of data 

```{r}
datadef <- datastand[dataset_clean$loan_status==1,]
vecmediannodef <- apply(datanodef, 2, median)
vecmediannodef

```

The absolute deviations of each element in the datade 
Subtracting the median values from the elements of datadef and then taking the absolute value of the differences.

```{r}

vecmediandef <- apply(datadef,2,median)
matabsdevnodef <- abs(datanodef - matrix(rep(vecmediannodef,nrow(datanodef)),nrow=nrow(datanodef),byrow=TRUE))
matabsdevdef <- abs(datadef - matrix(rep(vecmediandef,nrow(datadef)),nrow=nrow(datadef), byrow=TRUE))

matabsdevdef

```


```{r}
matabsdev.all <- rbind(matabsdevnodef, matabsdevdef)
matabsdev.all <- data.frame(dataset_clean$loan_status, matabsdev.all)

matabsdev.all
```

T-test for each independent variable 

```{r}

t.test(matabsdev.all$person_income[matabsdev.all$dataset_clean.loan_status == 0], matabsdev.all$person_income[matabsdev.all$dataset_clean.loan_status == 1], alternative="less",var.equal=TRUE)
```


```{r}
t.test(matabsdev.all$person_emp_length[matabsdev.all$dataset_clean.loan_status == 0], matabsdev.all$person_emp_length[matabsdev.all$dataset_clean.loan_status == 1], alternative="less",var.equal=TRUE)
```

```{r}
t.test(matabsdev.all$loan_amnt[matabsdev.all$dataset_clean.loan_status == 0], matabsdev.all$loan_amnt[matabsdev.all$dataset_clean.loan_status == 1], alternative="less",var.equal=TRUE)
```

```{r}
t.test(matabsdev.all$loan_percent_income[matabsdev.all$dataset_clean.loan_status == 0], matabsdev.all$loan_percent_income[matabsdev.all$dataset_clean.loan_status == 1], alternative="less",var.equal=TRUE)
```

Insight: 
p-value for personal_income is < 0.05. Thus, rejecting the null hypothesis. 

The mean absolute deviation of non-defaulted loans is less than that of defaulted loans. 

Statistical tests to compare the mean absolute deviations of certain variables between defaulted and non-defaulted loans, helping to understand if there are significant differences between the groups.


```{r}
#datastand
```

```{r}
#datastand.all <- data.frame(dataset_clean$loan_status,datastand)
#colnames(datastand.all) <- colnames(dataset_clean[3:8])
#t2test <- hotelling.test(person_income + person_emp_length + loan_amnt + loan_int_rate + loan_percent_income + cb_person_cred_hist_length ~ dataset_clean$loan_status, data = datastand.all )
#cat("T2 statistic =", t2test$stat[[1]],"\n")
#print(t2test)
```

```{r}
#datastand.all
```

