---
title: "Factor analysis"
output: html_document
date: "2024-03-24"
---

```{r}

# Factor Analysis

library(psych)
data <- read.csv("C:/Users/rusha/Downloads/Multivariate/credit_risk_dataset.csv/credit data .csv", header=TRUE, fill = TRUE)
data <- na.omit(data[,2:6])
str(data)
```


```{r}
data$person_home_ownership <- as.factor(data$person_home_ownership)
data$person_income <- as.numeric(data$person_income)
data$loan_amnt <- as.numeric(data$loan_amnt)
data$cb_person_cred_hist_length <- as.numeric(data$cb_person_cred_hist_length)
str(data)
```


```{r}
sampled_data <- data[sample(nrow(data), 70), ]
matstd.data <- scale(sampled_data[,-1])
```

Q. Decide how many Factors are ideal for your dataset?

```{r}
fa.parallel(sampled_data[-1])
```

Parallel analysis suggests that the number of factors =  3  and the number of components =  2

```{r}
fit.pc <- principal(sampled_data[-1], nfactors=3, rotate="varimax")
fit.pc
```

```{r}
round(fit.pc$values, 3)
fit.pc$loadings
```

Insights: 1.385 1.210 0.947 0.458 are the rounded eigenvalues. 
Larger eigenvalues suggest that the corresponding principal component explains more variance in the data.

Under each principal component, you have loadings for each original variable.
Higher absolute values of loadings indicate a stronger correlation between the variable and the principal component.

In short, this output provides information on the variance explained by each principal component(eigenvalues) and the relationship between original variables and these components in the PCA(loadings).


```{r}
fa.plot(fit.pc) # See Correlations within Factors

```

Q. Show the columns that go into each factor 

```{r}
fa.diagram(fit.pc)
```

Q. Perform some visualizations using the factors 

```{r}
vss(sampled_data[-1])
```

```{r}
biplot(fit.pc)

```




