---
title: "Loan data- Multiple regression"
output: html_document
date: "2024-04-15"
---
1. Model Development
```{r}
loan_data <- read.csv("C:/Users/rusha/Downloads/Multivariate/credit_risk_dataset.csv/credit_risk_dataset.csv")
str(loan_data)

#loan_data_numeric <- as.data.frame(sapply(loan_data, as.numeric))
model_fit <- lm(loan_status ~ person_income + loan_amnt + loan_int_rate + cb_person_cred_hist_length + cb_person_default_on_file, data = loan_data)
summary(model_fit)
coefficients(model_fit)

```

Intercept: indicates the estimated value of the dependent variable (loan_status) when all predictor variables are zero. 0.6091.

For a one-unit increase in person_income, the expected change in loan_status is a decrease of approximately 0.00000457.
Very low p-value (< 0.001), i.e has a significant effect. 

For a one-unit increase in loan_amnt, the expected change in loan_status is an increase of approximately 0.00001073. 
very low p-value (< 0.001), i.e significant effect. 

For a one-unit increase in loan_int_rate, the expected change in loan_status is an increase of approximately 0.02467. 
very low p-value (< 0.001), i.e significant effect. 

For a one-unit increase in cb_person_cred_hist_length, the expected change in loan_status is a decrease of approximately 0.03373. 
p-value for this coefficient is 0.107, i.e not statistically significant. 

For cb_person_default_on_file set to 'Y' (yes) results in an increase of approximately 0.00972 in the expected value of loan_status. 
p-value for this coefficient is 0.828, i.e not statistically significant. 

The Adjusted R-squared value is approximately 0.4764, indicating that the model explains about 47.64% of the variability in loan_status.
The F-statistic is 81.99 with a p-value < 0.001, indicating that the overall model is statistically significant.

loan_status has statistically significant predictors (person_income, loan_amnt, and loan_int_rate). However, some predictors (cb_person_cred_hist_length and cb_person_default_on_fileY) are not statistically significant. 


2. Model Acceptance
```{r}
# Assess the acceptance of the model
# Check R-squared and adjusted R-squared
r_squared <- summary(model_fit)$r.squared
adj_r_squared <- summary(model_fit)$adj.r.squared
cat("R-squared:", r_squared, "\n")
cat("Adjusted R-squared:", adj_r_squared, "\n")

# Check the F-statistic and its p-value
f_statistic <- summary(model_fit)$fstatistic
f_statistic_value <- f_statistic[1]
f_statistic_p_value <- pf(f_statistic_value, f_statistic[2], f_statistic[3], lower.tail = FALSE)
cat("F-statistic:", f_statistic_value, "\n")
cat("p-value:", f_statistic_p_value, "\n")
```
R-squared (0.4823) indicates that approximately 48.23% of the variability in the dependent variable (loan_status) is explained by the model.
Adjusted R-squared (0.4764) adjusts the R-squared value for the number of predictors in the model. It is slightly lower than R-squared, suggesting that adding more predictors may not significantly improve the model's explanatory power.

The F-statistic (81.98651) tests the overall significance of the model. It compares the fit of the model with no predictors (null model) against the fit of the current model.
The associated p-value (1.039549e-60) is extremely low, indicating strong evidence against the null hypothesis (that the model has no predictive power).

However,  the model appears to have some predictive power and explanatory value, there may be opportunities to refine the model by considering additional predictors or exploring alternative modeling techniques.


3. Residual Analysis 
```{r}
# Select only numeric variables from fish_data
numeric_loan_data <- loan_data[, sapply(loan_data, is.numeric)]
# Plot pairs
pairs(numeric_loan_data, main = "Loan Data")
confint(model_fit,level=0.95)
fitted(model_fit)
residuals(model_fit)

# Plot residuals vs. fitted values
plot(model_fit, which = 1)

# Plot normal Q-Q plot of residuals
plot(model_fit, which = 2)
```

We see the points are not always normally distributed. 

4. Prediction
```{r}
# Make predictions
predictions <- predict(model_fit, newdata = loan_data)

# Print predictions
print(predictions)
```

Each predicted weight represents the estimated weight of loan i.e if a person gets it or not, based on its corresponding values of the predictor variables (person_income + loan_amnt + loan_int_rate + cb_person_cred_hist_length + cb_person_default_on_file) as per the fitted regression model.

5. Model Accuracy
```{r}
# Calculate Mean Squared Error (MSE)
mse <- mean(model_fit$residuals^2)

# Calculate Root Mean Squared Error (RMSE)
rmse <- sqrt(mse)

# Calculate Mean Absolute Error (MAE)
mae <- mean(abs(model_fit$residuals))

# Print the accuracy metrics
cat("Mean Squared Error (MSE):", mse, "\n")
cat("Root Mean Squared Error (RMSE):", rmse, "\n")
cat("Mean Absolute Error (MAE):", mae, "\n")
```
The MSE is 0.127923, on average, the squared difference between the observed and predicted values. 

The RMSE is 0.3576632, suggesting that, on average, the difference between the actual and predicted values.

The MAE is 0.2890811, indicating that, on average, the absolute difference between the actual and predicted values. 

Lower values of MSE, RMSE, and MAE indicate better model performance, with smaller differences between actual and predicted values.


```{r}
#Anova Table
anova(model_fit)
vcov(model_fit)
cov2cor(vcov(model_fit))
temp <- influence.measures(model_fit)
temp
plot(model_fit)

anova_result <- anova(model_fit)
print(anova_result)
```

