---
title: "PCA analysis"
output: html_document
date: "2024-03-01"
---

```{r}
library(readr)
unclean_data <- read_csv("C:/Users/rusha/Downloads/Multivariate/credit_risk_dataset.csv/credit_risk_dataset.csv")
dataset <- na.omit(unclean_data)
var_data <- dataset[,2:8]
str(dataset)
```

```{r}
attach(var_data)
cor(var_data[-1])
```

```{r}
data_pca <- prcomp(var_data[,-1],scale=TRUE)
data_pca
```

```{r}
summary(data_pca)
```
1. Decide how many Principal Components (PCs) you want to keep and why (2 points)
Answer: I want to keep PC1 PC2 and PC3, here is why
PC1: highest standard deviation: captures the most variability in data: 29.2% of the total variance is captured
PC2: highest standard deviation: captures the most variability in data: 23.06% of the total variance is captured
PC3: lower amount than PC1 and PC2: captures 16.85% of the total variance: cumulative of PC1, PC2 and PC3 is 69.1%.

PC4 to PC6 captures significantly less proportions of total variance. 


```{r}
(eigen_data <- data_pca$sdev^2)
```
Insight: We consider retaining principal components with eigenvalues greater than 1.

```{r}
names(eigen_data) <- paste("PC",1:6,sep="")
eigen_data
```

```{r}
sumlambdas <- sum(eigen_data)
sumlambdas
```

```{r}
propvar <- eigen_data/sumlambdas
propvar
```

```{r}
cumvar_data <- cumsum(propvar)
cumvar_data
```

```{r}
matlambdas <- rbind(sqrt(eigen_data),propvar,cumvar_data)
rownames(matlambdas) <- c("Eigenvalues","Prop. variance","Cum. prop. variance")
round(matlambdas,4)
```

```{r}
summary(data_pca)
```
2. Explain the variate representation each PCs

```{r}
data_pca$rotation
```

```{r}
print(data_pca)
```
Insight: 
PC1 positively correlated with person_income person_emp_length and cb_person_cred_hist_length, thus PC1 captures information about the overall financial status, employment length, and credit history.
loan_percent_income has a strong negative correlation with PC1, thus higher loan amounts relative to income are associated with lower values of PC1.

Similarily, PC2 is positively correlated with cb_person_cred_hist_length, strong negative correlation with loan_ammt and negatively correlated with all the other columns. 
PC3 is positively correlated with person_emp_length, strong negative correlation with cb_person_cred_hist_length and negatively with all other columns. 

```{r}
data_pca$x
```

```{r}
#Identify the score by the loan status
datatyp_pca <- cbind(data.frame(loan_status),data_pca$x)
datatyp_pca
```

```{r}
# Means of scores for all the PC's classified by Survival status
tabmeansPC <- aggregate(datatyp_pca[,2:7],by=list(loan_status=dataset$loan_status),mean)
tabmeansPC
```

Mean scores of each principal component for default vs non-default groups.

```{r}
tabfmeans <- t(tabmeansPC[,-1])
tabfmeans
```

```{r}
colnames(tabfmeans) <- t(as.vector(tabmeansPC[1]$loan_status))
tabfmeans
```

```{r}
# Standard deviations of scores for all the PC's classified by Survival status
tabsdsPC <- aggregate(datatyp_pca[,2:7],by=list(loan_status=dataset$loan_status),sd)
tabfsds <- t(tabsdsPC[,-1])
colnames(tabfsds) <- t(as.vector(tabsdsPC[1]$loan_status))
tabfsds
```

```{r}
t.test(PC1~dataset$loan_status,data=datatyp_pca)
```

```{r}
t.test(PC2~dataset$loan_status,data=datatyp_pca)
```

```{r}
t.test(PC3~dataset$loan_status,data=datatyp_pca)

```

```{r}
t.test(PC4~dataset$loan_status,data=datatyp_pca)

```

```{r}
t.test(PC5~dataset$loan_status,data=datatyp_pca)
```

```{r}
t.test(PC6~dataset$loan_status,data=datatyp_pca)
```

Insights: 
Based on the p-values: PC1, PC5 and PC6 has significant difference in mean scores between the default and non-default groups. 
For PC1  non-default group (mean = 1.0913824)  default group (mean = -0.8811571)
PC5      non-default group (mean = 0.1446998)  default group (mean = -0.1168273)

while PC2, PC3 and PC4 has no significant differences. 

```{r}
## F ratio tests
var.test(PC1~dataset$loan_status,data=datatyp_pca)
```

```{r}
var.test(PC2~dataset$loan_status,data=datatyp_pca)

```

```{r}
var.test(PC3~dataset$loan_status,data=datatyp_pca)

```

```{r}
var.test(PC4~dataset$loan_status,data=datatyp_pca)

```

```{r}
var.test(PC5~dataset$loan_status,data=datatyp_pca)

```

```{r}
var.test(PC6~dataset$loan_status,data=datatyp_pca)

```

```{r}
# Plotting the scores for the first and second components
plot(datatyp_pca$PC1, datatyp_pca$PC2,pch=ifelse(datatyp_pca$loan_status == 0,1,16),xlab="PC1", ylab="PC2", main="Status of loan against values for PC1 & PC2")
abline(h=0)
abline(v=0)
legend("topright", legend=c("No-default","Default"), pch=c(1,16))
```

```{r}
plot(eigen_data, xlab = "Component number", ylab = "Component variance", type = "l", main = "Scree diagram")
```

```{r}
plot(log(eigen_data), xlab = "Component number",ylab = "log(Component variance)", type="l",main = "Log(eigenvalue) diagram")
```

```{r}
plot(dataset[,-1])
```

```{r}
plot(data_pca)
```

```{r}
library(factoextra)
fviz_eig(data_pca, addlabels = TRUE)
```


```{r}
library(FactoMineR)
library(ggfortify)
library(psych)
library(corrplot)
library(devtools)
pairs.panels(dataset[,-1],
             gap = 0,
             bg = c("red", "blue")[dataset$loan_status],
             pch=21)
```


```{r}
fviz_pca_var(data_pca,col.var = "cos2",
             gradient.cols = c("#FFCC00", "#CC9933", "#660033", "#330033"),
             repel = TRUE)
```


