---
title: "social media"
output: html_document
date: "2024-03-25"
---

```{r}
library(readr)
library(FactoMineR)
library(ggfortify)
library(psych)
library(corrplot)
library(devtools)
data <- read.csv("C:/Users/rusha/Downloads/Multivariate/mva class social media.csv")
str(data)
```

```{r}
#data$OTT <- data$OTT..Netflix..Hulu..Prime.video..
summary(data)
```


```{r}
stddata <- scale(data)
stddata
```


```{r}
datacov <- cov(data)
datacov

# Formatting 
datacov_rounded <- round(datacov, digits = 2)
datacov_df <- as.data.frame(datacov_rounded)
datacov_df
```

```{r}
datamean <-  colMeans(data)
datamean
```

```{r}
data_MD <- mahalanobis(data, datamean, datacov)
data_MD

#My usage vs class
data_MD[4]

```

```{r}
# Lets plot correlation between columns
cor.plot(data)
```

Insights: 
1) Instagram is highly correlated with Snapchat, Whatsapp/Wechat.
2) Tired waking up is correlated with Trouble falling asleep.
3) Trouble falling asleep is correlated with Instagram, Snapchat.
4) Tired waking up is correlated with Linkedin, Youtube but in opposite direction. 


```{r}
#Lets identify outliers in our data
boxplot(data)
```

## Principal Component Analysis

```{r}
cor(data)
```

```{r}
data_pca <- prcomp(data,scale. =TRUE)
data_pca
```

```{r}
summary(data_pca)
```

Insights : Keeping PC1, PC2, PC3 and PC4, as it captures ~75% of variance. 

## Checking if our insight is correct 

```{r}
(eigen_data <- data_pca$sdev^2)
```

```{r}
plot(eigen_data, xlab = "Component number", ylab = "Component variance", type = "l", main = "Scree diagram")

```
```{r}
plot(eigen_data, type = "b", main="Scree Plot")
abline(h=1,col="blue",lty=2)
```

```{r}
library(factoextra)
fviz_eig(data_pca, addlabels = TRUE)
```

```{r}

fviz_pca_var(data_pca,col.var = "cos2",
             gradient.cols = c("#FFCC00", "#CC9933", "#660033", "#330033"),
             repel = TRUE)
```

Plot shows which variables are correlated. Here, Linkedin, Snapchat, Instagram, Youtube, Whatsapp/Wechat, Mood Productivity.

## Clustering 
#1) Hierarchical

```{r}
library(cluster)
library(readr)
library(factoextra)
library(magrittr)
library(NbClust)

dist.data <- dist(stddata, method="euclidean")
colnames(dist.data) <- rownames(dist.data)

clusdata.nn <- hclust(dist.data, method = "single")

options(repr.plot.width=10, repr.plot.height=6)  # Adjust the plot size as needed
plot(as.dendrogram(clusdata.nn), ylab="Distance",
     main="Dendrogram of all Users")
```
Insight: I think the optimal number of clusters are 5 

## Non-hierarchical clustering/ K-means clustering

```{r}
(kmeans2.data <- kmeans(stddata,4,nstart = 10))

```

```{r}
gap_stat <- clusGap(stddata, FUN = kmeans, nstart = 10, K.max = 10, B = 50)

fviz_gap_stat(gap_stat)
```

Optimal number of cluster verified with the above plot. 

```{r}
set.seed(123)
## Perform K-means clustering
km.res3 <- kmeans(stddata, 3, nstart = 25)  
# Visualize clusters
fviz_cluster(km.res3, data = stddata,  
             ellipse.type = "convex",
             palette = "jco",
             ggtheme = theme_minimal())
```
From the plot we cannot directly infer anything. We see a lot of overlapping clusters. That is because we have a lot users that are using the same social media platforms. 

```{r}
set.seed(123)
## Perform K-means clustering
km.res5 <- kmeans(stddata, 5, nstart = 25)  
# Visualize clusters
fviz_cluster(km.res5, data = stddata,  
             ellipse.type = "convex",
             palette = "jco",
             ggtheme = theme_minimal())
```

```{r}
res.hc <- stddata %>% scale() %>% dist(method = "euclidean") %>%
  hclust(method = "ward.D2")  # Change matstd.data to your dataset

# Visualize the Dendrogram
fviz_dend(res.hc, k = 5,  # Cut in four groups
          cex = 0.5,  # label size
          k_colors = c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07", "pink"),
          color_labels_by_k = TRUE,  # color labels by groups
          rect = TRUE)
```
## Factor Analysis
```{r}
fa.parallel(data[-1])
```

```{r}
fit.pc <- principal(data[-1], nfactors=3, rotate="varimax")
fit.pc
```

```{r}
fa.diagram(fit.pc)
```

```{r}
vss(data[-1])
```

