---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r}
social_data <- read.csv("C:/Users/rusha/Downloads/Multivariate/mva class social media.csv")
str(data)
```

```{r}
model_fit <- lm(Mood.Productivity ~ Instagram + Whatsapp..Wechat + Youtube + Linkedin, data = social_data)
summary(model_fit)
coefficients(model_fit)
```
Intercept: 
FOr one unit increase in (Instagram/Whatsapp..Wechat/Linkedin), the Mood/productivity of a person increases by (0.012/0.013/0.032) respectively. And for one unit increase in YouTube, the Mood decreases by 0.04.

From the Adjusted R-squared value, the model only explains an 8.9% of variance in data. 
This is not the best model. 

2. Model Acceptance
```{r}
# Assess the acceptance of the model
# Check R-squared and adjusted R-squared
r_squared <- summary(model_fit)$r.squared
adj_r_squared <- summary(model_fit)$adj.r.squared
cat("R-squared:", r_squared, "\n")
cat("Adjusted R-squared:", adj_r_squared, "\n")

# Check the F-statistic and its p-value
f_statistic <- summary(model_fit)$fstatistic
f_statistic_value <- f_statistic[1]
f_statistic_p_value <- pf(f_statistic_value, f_statistic[2], f_statistic[3], lower.tail = FALSE)
cat("F-statistic:", f_statistic_value, "\n")
cat("p-value:", f_statistic_p_value, "\n")
```

Although, R-squared: 0.271 suggests the model explains moderate amount of variance, the F-statistic is 1.49 and p-value is 0.2511, indicating that the model is not significant.

3. Residual Analysis 
```{r}
# Select only numeric variables from fish_data
numeric_social_data <- social_data[, sapply(social_data, is.numeric)]
# Plot pairs
pairs(numeric_social_data, main = "Loan Data")
confint(model_fit,level=0.95)
fitted(model_fit)
residuals(model_fit)

# Plot residuals vs. fitted values
plot(model_fit, which = 1)

# Plot normal Q-Q plot of residuals
plot(model_fit, which = 2)
```

Most of the points lies on the straight line, thus the points are normally distributed.

4. Prediction
```{r}
# Make predictions
predictions <- predict(model_fit, newdata = social_data)

# Print predictions
print(predictions)
```

5. Model Accuracy
```{r}
# Calculate Mean Squared Error (MSE)
mse <- mean(model_fit$residuals^2)

# Calculate Root Mean Squared Error (RMSE)
rmse <- sqrt(mse)

# Calculate Mean Absolute Error (MAE)
mae <- mean(abs(model_fit$residuals))

# Print the accuracy metrics
cat("Mean Squared Error (MSE):", mse, "\n")
cat("Root Mean Squared Error (RMSE):", rmse, "\n")
cat("Mean Absolute Error (MAE):", mae, "\n")
```

Lower values of MSE, RMSE, and MAE indicate better model performance, with smaller differences between actual and predicted values.


```{r}
#Anova Table
anova(model_fit)
vcov(model_fit)
cov2cor(vcov(model_fit))
temp <- influence.measures(model_fit)
temp
plot(model_fit)

anova_result <- anova(model_fit)
print(anova_result)
```

